\LARGE \bf Explicit Knowledge and the Deliberative Layer: Lessons Learned

### 2013-03-21 paper.tex
Added cognitive laod figure
Large proofreading + section on Theory of Mind

\usepackage[utf8x]{inputenc}

@@

Natural interaction and cooperation are the current (dare we say,
\emph{short-term}) targets for the human-robot interaction community. They are
keys to the broad class of \emph{interactive manipulation problems}: several
agents agree on a (more or less implicit) joint goal that requires some sort of
cooperation to be successfully achieved. This class of problems involves both
dialogue and manipulation and they are often not completely defined at
start-up: they require iterative and interactive resolution.

Over the last years we have focused our efforts on identifying the cognitive
prerequisites of these challenges, and giving them experimental reality on the
robots: what is required for a sentence like ``Let's dress the table together''
to be understood by the robot, correctly interpreted in the spatial and
temporal context of the interaction, and eventually transformed into a set of
actions?

@@

We focus in this paper on that last item, to present what that means, and what
are the benefits and drawbacks.

@@

While we do not present them here, the ideas and techniques pertaining to this
paper have been all implemented and tested on several robots (see previously
reported experiments in~\cite{Lemaignan2010, Ros2010b, Lemaignan2011a,
Warnier2012a}). Videos presenting some of these experimental results can also
be watched online at~\url{http://www.laas.fr/~slemaign}.

@@

We first describe with more details how the deliberative layer of our robots is
built and why this monolithic functional group enables what we call a
\emph{knowledge-oriented} architecture. The first section also introduce our
central knowledge management tool, the ORO server.

The next sections go over each of the facets of this deliberative layer: how
knowledge is produced asynchronously from geometric reasoning, how the robot
conducts grounded multi-modal interaction with humans, how knowledge is used by
decision-making components like the robot controller or the task planner, and
finally, how our knowledge-centric architecture enables the implementation of
specific internal cognitive processes.

We conclude the article with a discussion about the strenghts and weaknesses of
such a knowledge-oriented architecture, fron three perspectives: the architect,
the IA expert and the cognitician.

@@

\cite{Alami2011}

\begin{figure*}
        \centering
        \includegraphics[width=1.7\columnwidth]{archi}
        \caption{Overview of the LAAS deliberative layer.}
        \label{fig|archi}
\end{figure*}

We can give a broader look at the knowledge and the streams of knowledge in our
systems.  Based on the experience gained while developing and deploying {\sc
ORO}, our ontology-based knowledge server, we have presented how symbolic
knowledge could be produced from perception and geometric reasoning in modules
like {\sc SPARK}, a grounded, perspective-aware, geometric reasoner. We have
seen how symbolic knowledge could be reused by different control systems and
task planners like {\sc Cram}, {\sc SHARY}, {\sc pyRobots}, the {\sc CLSU
Toolkit} or {\sc HATP} and how they take advantage of semantic abstractions
provided by knowledge base. We have also presented the bidirectional
integration of {\sc Dialogs}, a natural language processor for English, with
the knowledge base.

Altogether, these components compose an architecture that we call
\emph{knowledge-oriented}:

@@

    \item{and a \emph{symbolic} reactive, event-driven approach to supervision.
    By managing events at the same level as the reasoner, we take full
    advantage of the inference abilities of ORO to trigger events whose
    \texttt{true} conditions can be inferred.}

@@

Regarding the anchoring question, this architecture is bidirectional. The
components we described provide a \textit{bottom-up} grounding process: SPARK
and \textsc{Dialogs} constantly build and push new symbolic contents about the
world to ORO where it becomes accessible to decisional layers. In parallel, ORO
relies on reasoning in a \textit{top-down} way to produce new facts that may
trigger in return physical behaviours.

@@

This architecture may also contribute to bring closer robotics and psychology:
it provides clear entry points to implement some classical psychology tests to
robots. For instance, we presented experiments focused on issues related to
perspective taking. By explicitly enabling independent modeling of the beliefs
of each agent, our architecture is especially well suited to set up cognitive
and psychological experiments that involve a theory of mind, such as
\emph{False-Belief} experiments, as recently presented in~\cite{Warnier2012a}.

\subsection{Knowledge Model}

\cite{Lemaignan2010}

\begin{figure}
        \centering
        \includegraphics[width=\columnwidth]{spark-oro}
        \caption{Functional overview of knowledge base (\emph{oro-server}, top part) and the geometric situation assessment module (\emph{SPARK}, bottom part)}
        \label{fig|spark-oro}
\end{figure}

One of the field of human-robot interaction where the introduction of the
semantic layer has been most beneficial is natural language processing.
\cite{Lemaignan2011a} details the techniques and the tool called
\texttt{dialogs} that we have developed for English language parsing and
grounding, for (minimalist) dialogue management and for verbalisation.

Natural language input (in experiments, we rely on an Android-based interface
running on a tablet, with Google-based speech recognition, for the input) is
parsed into a grammatical structure, and each sentence atoms are resolved with
the help of the ontology to ground classes (when a user says ``pick the can'',
to which instance of \emph{can} the user refers to?) and actions. Sentences are
sorted into questions, desires and statements, and processed accordingly.

The system supports quantification (``give me \{a|the|some|all|any|...\} can''),
thematic roles attached to actions, interactive disambiguation (the robot asks
questions when needed), anaphora resolution (``give \emph{it} to me'') based on
dialogue history. It also permits knowledge extension by learning new semantic
structures (for instance, a sentence like ``learn that cats are animals'' is
converted into \stmt{Cat subClassOf Animal}, even if \concept{Cat} is not known
beforehand) and tries to translate \emph{states} (``I'm tired'') into
\emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1 hasFeature
tired}).

\subsection{Disambiguation at semantic level}

One prototypical example of semantic disambiguation has been given in
\cite{Ros2010b} with the child's \emph{spygame}: two players are facing
each other with a set of random objects in-between, one player mentally choose
one object, and the other player has to guess the object by asking closed
questions like \emph{Is your object small or large?}. Based on the knowledge it
has acquired, the robot is able to minimize the number of questions required to
find the object.

When playing this kind of game, however, the issue arises that the robot has no
way to select which knowledge about the object is relevant in the interaction
context. For instance, the knowledge base may store facts like \stmt{obj1 type
ActiveConcept} (which internally means that this concept was mentioned in a
discussion in the last few seconds): this information is not a relevant
property of \concept{obj1} when trying to disambiguate concepts with humans.
This distinction between \emph{internal knowledge} (meaningful to
the systme only) and \emph{common knowledge} (whose meaning is understood by
all the interactors) has not been properly dealt with in our architecture.

Besides, even knowledge that belongs to what we call \emph{common knowledge},
may not be appropriate in a given interaction context. For instance, the system
may compute that at a given instant, the human is looking at the object:
\stmt{human looksAt obj1}. This property makes sense to both parties, but in
the context of the \emph{spygame}, we would like to mainly use immanent
properties, not volatile like a gaze. More research is required to precise this
idea of contexts and knowledge class attached to them.

\fxnote{Mention failures like 'placeOing'}

Because the main communication channel between component is semantic data,
different communication modalities (explicit like verbal, dectic or based on
gaze, or implicit like postures) are accessed in an homogeneous way.

The grounding process copes with and integrates the different physical
modalities at two disctint levels.

First, the grounding module explicitely check for the presence and value of
particular facts: for instance, when a set of instances match a category (the
human says ``give me the bottle'' and the robot knows about three bottles), the
module may decide (it actually depends on the quantifier preceding the class:
\emph{the}, \emph{a}, \emph{any}...) to discard some of them based on their
\emph{visibility} by the speaker (implicit communication context built from the
human posture).

Another example, when the human says ``this'', the robot checks if the human is
currently both pointing and looking at some object. In that case, \emph{this}
is replaced by the object focused on.

Note that, while the system benefits from the complementary modalities, they
are not required. The dialogue system would work as well in a purely verbal
setting.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robot Control}
\label{sect|ctrl}

In order to devise how a given goal can be accomplished, the robot has to
elaborate a plan,\ie a set of actions to be achieved by the robot and its human
partners.  We use in our architecture the HATP planner~\cite{Alili2009} (for
Human Aware Task Planner).  HATP is based on a Hierarchical Task Network (HTN)
refinement, which performs an iterative task decomposition into sub-tasks until
reaching atomic actions.  The planning domain defines a set of methods
describing how to decompose a task and can be seen as the procedural knowledge
of the robot.

In order to produce a collaborative behaviour, the robot plans not only for
itself but also for the other agents. The planning domain of each agent is
instantiated from the agent-specific model in the ORO server. The resulting
plan, called ``shared plan'' is a set of actions that form a stream for each
agent involved in the goal achievement. HATP can also be tuned by setting up
different costs depending on the actions to apply and by taking into account a
set of constraints called \emph{social rules}. This tuning aims at adapting the
robot's behaviour according to the desired level of cooperation of the robot.

Second remark, after some research (see appendix B of~\cite{ma these} for a
detailled discussion)  we have decided not to represent neither tasks nor plans
in the ontology: the planning domain (with tasks pre- and postconditions) is
stored in a specific format, outside of the knowledge repository, and the
resulting plans are directly communicated to the robot controller. The ontology
formalism that we have adopted (namely OWL) does not appear to be a satisfaying
reprepresentation language for concepts that are both tightly bound to time
representation and non-monotonic. Other researches (\cite{tenorth...}) have attempted with more success (\fixme{regarder ca!!}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Internal Cognitive Processes}
\label{sect|intern}

\subsection{Theory of Mind}

\cite{Warnier2012a}

\subsection{Working Memory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sect|conclusion}

We conclude this article with three different perspectives on the knowledge-centric
techniques we have presented here: first, the \emph{architect perspective}:
what are the pro and cons of our approach from an engineer point of view. Then,
the \emph{logician perspective}: does the promise of automated reasoning for
robots really comes to fruition with ontologies? And eventually the
\emph{cognitician perspective} or how explicit knowledge manipulation endows
our robots with new cognitive capabilities.

\subsection{The Architect view: loose coupling and modalities merging}

@@

Since these statements are anchored in the robot's pool of knowledge, they
actually convey \emph{meaning} through the system (for instance, when the
dialog module or the task planner or the execution controller uses the symbol
{\tt Give} \emph{means}, it refers to the abstract class of all give actions, which are also stated
to be a kind of purposeful actions). This has one important consequence from
the perspective of the system architecture: modules interfaces are only defined
in term of knowledge consumption and/or production. This brings good decoupling
properties. For instance, the output of the dialog module is a new
\emph{situation desired} by the speaker. This module can be transparently
replaced or completed by any other module that produces the same chuncks of
knowledge. Only a thin protocol layer remains to get the modules to interact
together. This allows us to implement (otherwise difficult) merging of
radically different perception modalities like vision (the robot sees an object on
a table) and verbal description (a human says to the robot another object is
under the table): both are converted to RDF statements ({\tt obj1 type Object},
{\tt obj1 isOn table1} and {\tt obj2 type Object}, {\tt obj2 isUnder table1})
and made available to the other robot modules.

@@

And obviously, many aspects like uncertainty management, modal logics,
non-monotonic reasoning,... remain to be explored to gain the level of
expressiveness required to cover many of the more complex human-robot
interaction scenarii.

\subsection{The Logician view: the importance of trivial inferences}

@@

This example of reasoning, quoted from Moritz Tenorth, is a good example of
simple yet non-trivial reasoning. As a matter of fact, only very few of such
reasoning cases where positively identified in our scenarii and experiments
(and consequently implemented as rules in ORO).

The design choices of our architecture partially explain that fact: first, the
planning task (which is the prototypical reasoning task) is delegated to a
dedicated, external planner. Then, time is not represented in ORO, and
consequently no temporal reasoning takes place at this level: action
recognition or monitoring are handled by other layers, and the underlying
reasoning tasks are not implemented as explicit symbolic rules in the knowledge
base.

The experiments we have conducted are also likely to have too simplistic
semantics to let complex reasoning needs to emerge. Scenarii with more complex
semantics would be desirable to better stress the expressiveness and inference
abilities provided by description logics.

Is reasoning at the knowledge level immature or even superfluous, then? Not so:
hundreds of trivial (from a human point of view) inferences are continuously
produced by the system (translating inheritance relations, domain/range
constraints, transitivity, etc.) and encode a large amount of common-sense
knowledge that would be tedious, to say the least, to manually assert. These
trivial inferences are all the more important that an expressive knowledge
representation language is used: when a language like OWL allows to directly
represent high-level concepts like partitions, cardinality restrictions,
properties' ranges and domains, it leads to a more implicit (because more
abstract) description of the vocabulary that in turn requires more underlying
reasoning. With the progress in the understanding of the relations between
expressiveness and (tractable) satisfiability, along with the progress of
reasoners, more and more of the inferences do not need to be explicit anymore,
and consequently move behind the scenes.

And we predict that \emph{common-sense encoding} is likely to remain the main
application of reasoning in our robotic architectures, where reasoning related
to decision making mostly happen outside the knowledge representation system.

\subsection{The Cognician view: palpable knowledge and semantic thinking}

The main motive goal was to transform the knowledge in the robot from some ubiquitous,
pervasive, multi-modal and, most importantly, mostly undefined feature of the
system into an observable, quantifiable, manipulable resource, what we could
call a \emph{palpable} feature.

This transformation, both from the technical point of view (the ORO server, the
ontologies, the bindings, etc.), and as a more subtle change in the practises
related to the development of robotic components, is the main contribution of
this thesis.

@@

Communication between the robot's module is now conceived in term of what are
the \emph{semantics} of the information flows, instead of a simple
compatibility of interfaces. When defining the frontiers of a robotic
component, we do not think anymore only in term of ``is the interface complete
and self-contained'', but also in term of ``is the semantic complete and
consistent?''. This allow a deeper, more correct modularity: two modules that
share the same, well-defined semantic can be confidently exchanged. When we
remove or disable a component (the dialogue processing, the geometric
reasoning, ...), we know precisely what knowledge will not be available
anymore.

@@

>>>>

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

@@

Natural interaction and cooperation are on the (dare we say,
\emph{short-term}) agenda for the human-robot interaction community. They are
keys to the broad class of \emph{interactive manipulation problems}: several
agents agree on a (more or less implicit) joint goal that requires some sort of
cooperation to be successfully achieved. This class of problems involves both
dialogue and manipulation and they are often initially underspecified: they
require iterative and interactive resolution.

Over the last years we have focused our efforts on identifying the cognitive
prerequisites of these challenges, and giving them experimental reality on the
robots: what is required for sentences like ``Let's dress the table together''
to be understood by the robot, correctly interpreted in the spatial and
temporal context of the interaction, and eventually transformed into a set of
actions.

@@

The last item is the focus of this paper: what does ``knowledge-centric''
architecture means, and what are its benefits and drawbacks.

@@

While we do not present experiments in this paper, the ideas and techniques we
present all have been implemented and tested on several robots (see previously
reported experiments in~\cite{Lemaignan2010, Ros2010b, Lemaignan2011a,
Warnier2012a}). Also, videos presenting some of these experimental results can
be watched online at~\url{http://www.laas.fr/~slemaign}.

@@

In the next section, we first introduce the deliberative layer of our robots at
the functional level. It represents what we call a \emph{knowledge-oriented}
architecture, built around a central knowledge management tool called the {\sc
Oro} server.

The following sections go over each of the facets of this deliberative layer:
how knowledge is produced asynchronously from geometric reasoning, how the
robot conducts grounded multi-modal interaction with humans, how knowledge is
used by decision-making components like the robot controller or the task
planner, and finally, how our knowledge-centric architecture enables the
implementation of specific internal cognitive processes.

We conclude the article with a discussion about the strengths and weaknesses of
such a knowledge-oriented architecture, from three perspectives: the software
architect, the IA expert and the cognitician.

@@

\cite{Alami2011}
\ref{fig|archi}

\begin{figure*}
        \centering
        \includegraphics[width=1.7\columnwidth]{archi}
        \caption{Overview of the LAAS deliberative layer. Knowledge is
        centrally managed in an active \emph{semantic blackboard}, pictured
        above with a thick border.}
        \label{fig|archi}
\end{figure*}

Regarding the anchoring question, this architecture is bidirectional. The
components we described provide a \textit{bottom-up} grounding process: SPARK
and \textsc{Dialogs} constantly build and push new symbolic contents about the
world to {\sc Oro} where it becomes accessible to decisional layers. In parallel, {\sc Oro}
relies on reasoning in a \textit{top-down} way to produce new facts that may
trigger in return physical behaviours.

\subsection{Knowledge model}

In our architecture (Fig.~\ref{fig|archi}), knowledge manipulation relies on a
\emph{semantic blackboard}: one central server (the {\sc Oro}
server~\cite{Lemaignan2010}) stores knowledge as it is produced by each of the
deliberative components. It conversely exposes an API to query the knowledge
base.

Knowledge is represented as RDF triples in the OWL sub-language. Each
time triples are added or removed from the knowledge base, a Description
Logics reasoner ({\sc Pellet}\footnote{\url{http://clarkparsia.com/pellet/}})
classifies the whole ontology and inserts all possible inferred triples.

Relying on RDF triples and Description Logics has major advantages like the
availability of numerous open-source libraries to manipulate the ontology,
interoperability with several major on-line knowledge bases (like {\sc
OpenCyc}, {\sc WordNet} or {\sc DBPedia}), open-world reasoning, and the formal
guarantee of decidability (it is always possible to classify a Description
Logics ontology).

It has also notable limitations: RDF triples imply only binary predicates
(\stmt{subject predicate object}), which constrains the expressiveness of the
system or leads to cumbersome reifications. Alternative exists (like {\sc
KnowRob}~\cite{Tenorth2009a}) that mixes RDF with more expressive logic
languages like {\sc Prolog}, at the price, however, of other limitations, like
closed-world reasoning or immutable T-Box. The classification
performance is another issue: from our experience, with an ontology sized for a
standard experiment (about 100 classes and 200 instances), classification
typically take about 100ms, which becomes problematic during interactions.
Besides, the performances are difficult to predict, since a seemingly
inoffensive new statement may indirectly change radically the logical
complexity of the whole knowledge model and lead to notable degradation of
classification time.

This knowledge model also largely exclude representation of continuous
phenomena (like time) or uncertain phenomena. When required, these have to be
managed by dedicated components.

Amongst the features of {\sc Oro}, the system can dynamically creates
new independent knowledge models (\ie independent instantiations of the
ontology). This feature is used for instance to store separately the beliefs of
each of the agents the robot interacts with (see section~\ref{sect|tom}).

\begin{figure}
        \centering
        \includegraphics[width=\columnwidth]{spark-oro}
    \caption{Functional overview of knowledge base ({\sc Oro} server, top part) and the geometric situation assessment module (\emph{SPARK}, bottom part)}
        \label{fig|spark-oro}
\end{figure}

\subsection{Ego-centric and allo-centric frames}

Natural language processing is one of the fields of human-robot interaction
where the introduction of the semantic layer has been most beneficial.
We detail in~\cite{Lemaignan2011a} the techniques and the tool called
\texttt{dialogs} that we have developed for English language parsing and
grounding, along with verbalisation and (minimalist) dialogue management.

Natural language input (in experiments, we rely on an Android-based interface,
with Google speech recognition) is parsed into a grammatical structure, and
each sentence atoms are resolved with the help of the ontology to ground
concepts like objects (\ie when a user says ``pick the can'', resolve to which
instance of \emph{can} the user is referring to) and actions. Sentences are
sorted into questions, desires and statements, and processed accordingly.

The system supports quantification (``give me \{a | the | some | all | any | ...\} can''),
thematic roles (action-specific predicates that qualify the actions), interactive disambiguation (the robot asks
questions when it needs more information), anaphora resolution (``give \emph{it} to me'') based on
dialogue history. It also permits knowledge extension by learning new semantic
structures (for instance, a sentence like ``learn that cats are animals'' is
converted into \stmt{Cat subClassOf Animal}, interprets common temporal and place adverbs (like \emph{above} or \emph{tomorrow}) and translates to a certain extend \emph{states} (``I'm tired'') into
\emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1 hasFeature
tired}).

\subsection{Limits of disambiguation at semantic level}

One prototypical example of semantic disambiguation has been given in
\cite{Ros2010b} with the child game \emph{spygame}: two players are facing
each other with a set of random objects in-between, one player mentally choose
one object, and the other player has to guess the object by asking closed
questions like \emph{Is your object small or large?} Based on the knowledge it
has acquired, the robot is able to minimize the number of questions required to
find the object.

When playing this kind of game, however, the issue arises that the robot has no
way to select which knowledge about the object is relevant in the interaction
context. For instance, the knowledge base may store facts like \stmt{obj1 type
ActiveConcept} (which internally means that this concept was mentioned in a
discussion in the last few seconds): this information is not a relevant
property of \concept{obj1} when trying to disambiguate concepts with humans.
This distinction between \emph{internal knowledge} (meaningful to
the system only) and \emph{common knowledge} (whose meaning is understood by
all the interactors) has not been properly dealt with in our architecture.

Besides, even knowledge that belongs the \emph{common knowledge} may not be
appropriate in a given interaction context. For instance, the system may
compute that at a given instant the human is looking at the object: \stmt{human
looksAt obj1}. This property makes sense to both parties, but in the context of
the \emph{spygame}, we would like to mainly use immanent properties, not
volatile like a gaze. More research is required to indentify relevant
interaction contexts and knowledge classes attached to them.

Because all components rely on the same RDF formalism to format their outputs,
the different communication modalities (\emph{explicit} like verbal, deictic or
based on gaze, or \emph{implicit} like postures) are presented in an
homogeneous way.

The dialogue grounding process makes use of them at two distinct levels.

First, particular steps of the grounding process explicitly check for the
presence and value of specific facts: for instance, when a set of instances
match a category (the human says ``give me the bottle'' and the robot knows
about three bottles), the module may decide (it actually depends on the
quantifier preceding the class) to discard some of them based on their
\emph{visibility} by the speaker (implicit communication context built on the
human posture).

Another example, when the human says ``this'', the robot checks if the human is
currently pointing at some object. In that case, \emph{this} is replaced by the
object focused on.

Note that, while the system benefits from the complementary modalities, they
are not all required. The dialogue system would for instance run equally well
with only the verbal modality.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Robot control}
\label{sect|ctrl}

\subsubsection{Event-driven control}

In order to devise how a given goal can be accomplished, the robot has to
elaborate a plan, \ie a set of actions to be achieved by the robot and its
human partner(s).  We use in our architecture the HATP planner (for Human Aware
Task Planner~\cite{Alili2009}).  HATP is based on a Hierarchical Task Network
(HTN) refinement, which performs an iterative task decomposition into sub-tasks
until reaching atomic actions.  The \emph{planning domain} defines the set of
methods describing how to decompose a task and can be seen as the procedural
knowledge of the robot.

In order to produce a collaborative behaviour, the robot plans not only for
itself but also for the other agents. The planning domain of each agent is
instantiated from the agent-specific model in the {\sc Oro} server. The resulting
plan, called \emph{shared plan} is a set of action streams, one per
agent involved in the goal achievement. HATP can also be tuned by setting up
different costs depending on the actions to apply and by taking into account a
set of constraints called \emph{social rules}. This tuning aims at adapting the
robot's behaviour according to the desired level of cooperation of the robot.

Second remark, after some research (see appendix B of~\cite{Lemaignan2012a} for
a detailed discussion)  we have decided not to represent neither the planning
domain nor the resulting plans in the knowledge base: the planning domain (with
tasks pre- and postconditions) is stored in a specific format, outside of the
knowledge repository, and the plans are directly communicated to the robot
controller. From our experience, the OWL formalism is not a satisfying
representation language for concepts that are both tightly bound to time
representation and non-monotonic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Internal cognitive processes}
\label{sect|intern}

\subsection{Cognitive load}

\begin{figure}
        \centering
        \includegraphics[width=\columnwidth]{figs/cognitive_load.png}
        \caption{This diagram shows the average number of interactions with the
        knowledge base during a one-hour long experiment with the robot.
        Interactions are either knowledge alteration (addition/removal of
        statements) or queries. They are averaged on a sliding window of 60 seconds.}
        \label{fig|cognitiveload}
\end{figure}

Fig.~\ref{fig|cognitiveload} is a plot of the log of interactions between the
robot modules and the knowledge base during a one-hour long experiment. Because
most of the communication between modules goes in fact through the knowledge
base in our architecture (Fig.~\ref{fig|archi}), it actually pictures the
intensity of internal information exchanges between the different deliberative
component of the robot over time. Since communication actually relies on
grounded semantic information, we can call this level of interaction the
\emph{cognitive load} of the robot.

This simplistic measurement needs to be refined to take into account the fact
that the raw number of RPC calls to the knowledge base only imperfectly
reflects the actual cognitive processes running on, but it opens interesting
perspective nevertheless. It would be interesting for example to compare
and look for correlations with the cognitive load of a human partner (measured
via EEG of qualitative feedback).

\subsection{Theory of Mind}
\label{sect|tom}

Theory of Mind (originally defined in~\cite{Premack1978}) is the cognitive
ability that a subject possesses to represent the mental state of another
agent, possibly including knowledge that contradicts the subject own model: for
example, a book can be at the same time \emph{visible} by myself, and \emph{not
visible} by you.

This skill, which is essential to understand others' perspectives during
interactions, appears in children around their third year. It supposes the ability to build, store and retrieve separate models of the knowledge of the interactors.

Our knowledge base implements such a mechanism: when the robots infers a new
agent has been introduced in the knowledge base, it initializes a new,
independent, ontology for this agent. All the ontologies that are created share
the same common-sense knowledge, but rely on each agent's perspective for the
actual instantiation: the robot (geometrically) computes that the book is in
its own field of view, but not in the human one. The robot knowledge contains
the fact \stmt{book isVisible true} while the human model contains \stmt{book
isVisible false}.

The \cite{Warnier2012a}

\subsection{Working memory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{sect|conclusion}

\subsection{Knowledge-oriented architecture}

Altogether, the components we have presented compose an architecture that we
call \emph{knowledge-oriented}:

@@

    \item{and a \emph{symbolic}, reactive, event-driven approach to robot control.
    By managing events at the same level as the reasoner, we take full
    advantage of the inference abilities of {\sc Oro} to trigger events whose
    \texttt{true} conditions can be inferred.}

@@

@@

\subsection{Three perspectives on knowledge manipulation for robots}

We conclude this article with three different perspectives on the knowledge-centric
techniques we have presented here: first, the \emph{architect perspective}:
what are the pro and cons of our approach from an engineer point of view. Then,
the \emph{logician perspective}: does the promise of automated reasoning for
robots really comes to fruition with ontologies? And finally the
\emph{cognitician perspective} or how explicit knowledge manipulation endows
our robots with new cognitive capabilities.

\subsubsection{The Architect view: loose coupling and modalities merging}

@@

Since these statements are anchored in the robot's pool of knowledge, they
actually convey shared and unambiguous \emph{meaning} through the system. This
has one important consequence from the perspective of the system architecture:
modules interfaces are only defined in term of knowledge consumption and/or
production. This brings good decoupling properties. For instance, the output of
the dialog module is a new \emph{situation desired} by the speaker. This module
can be transparently replaced or completed by any other module that produces
the same chuncks of knowledge. Only a thin protocol layer remains to get the
modules to interact together. This allows us to implement (otherwise difficult)
merging of radically different perception modalities like vision (the robot
sees an object on a table: the output would look like \stmt{obj1 isOn table1})
and verbal description (a human says to the robot the object on the table
is red: it leads to a query \stmt{? isOn table1} and then a new assertion:
\stmt{obj1 hasColor red}).

@@

On the sensing side also we have in certain cases to bypass the semantic layer
(extension of the symbolic task planner with geometric constraints is one
example).

And obviously, many aspects like uncertainty management, modal logics,
non-monotonic reasoning,... are not well addressed with the Description Logic
formalism, and remain to be explored to gain the level of expressiveness
required to cover many of the more complex human-robot interaction scenarii.

\subsubsection{The Logician view: the importance of trivial inferences}

@@

This example of reasoning, quoted from Moritz Tenorth, is a good example of
simple yet non-trivial reasoning. As a matter of fact, only very few of such
prototypical reasoning cases where positively identified in our scenarii and experiments
(and consequently implemented as rules in {\sc Oro}).

The design choices of our architecture partially explain that fact: first, the
planning task (which is a typical reasoning task) is delegated to a
dedicated, external planner (HATP). Then, time is not represented in {\sc Oro}, and
consequently no temporal reasoning takes place at this level: action
recognition or monitoring are handled by other layers, and the underlying
reasoning tasks are not implemented as explicit symbolic rules in the knowledge
base.

The experiments we have conducted are also likely to have too simplistic
semantics to let complex reasoning needs to emerge. Scenarii with more complex
semantics would be desirable to better stress the expressiveness and inference
abilities provided by Description Logics.

Is reasoning at the knowledge level immature or even superfluous, then? Not so:
hundreds of trivial (from a human point of view) inferences are continuously
produced by the system (translating inheritance relations, domain/range
constraints, transitivity, etc.) and encode a large amount of common-sense
knowledge that can not be otherwise asserted. These trivial inferences are all
the more important that an expressive knowledge representation language is
used: when a language like OWL allows to directly represent high-level concepts
like partitions, cardinality restrictions, properties' ranges and domains, it
leads to a more implicit (because more abstract) description of the vocabulary
that in turn requires more underlying reasoning. With the progress in the
understanding of the relations between expressiveness and (tractable)
satisfiability, along with the progress of reasoners, more and more of the
inferences do not need to be explicit anymore, and consequently are delegated
to the reasoner.

We think that \emph{common-sense encoding} is likely to remain the main
application of reasoning in our architecture, where reasoning related
to decision making mostly happen outside the knowledge representation system.

\subsubsection{The Cognician view: palpable knowledge and semantic thinking}

The original motivation to introduce explicit knowledge management in our
architecture was to transform the knowledge in the robot from some ubiquitous,
pervasive, multi-modal and, most importantly, mostly undefined feature of the
system into an observable, quantifiable, manipulable resource: a
\emph{palpable} feature of the system.

This transformation, both from the technical point of view (the {\sc Oro}
server, the ontologies, the bindings, etc.), and as a more subtle change in the
practises related to the development of robotic components, is probably the
main outcome of this work.

@@

Communication between the robot's module is now conceived in term of what are
the \emph{semantics} of the information flows, instead of a simple
compatibility of interfaces. When defining the frontiers of a robotic
component, we do not think anymore only in term of \emph{is the interface complete
and self-contained}, but also in term of \emph{is the semantic complete and
consistent?}. This allow a deeper, more correct modularity: two modules that
share the same, well-defined semantic can be confidently exchanged. When we
remove or disable a component (the dialogue processing, the geometric
reasoning, ...), we know precisely what knowledge will not be available
anymore.

@@

Even more, this architecture also contribute to bring closer robotics and
cognitive psychology: it provides clear entry points to implement some
classical psychology tests to robots, be it related to perspective taking, to
language understanding, to a theory of mind (such as the \emph{False-Beliefs}
experiment), etc.
